---
title: "PML - Prediction Assigment"
author: "SK"
date: "July 26, 2014"
output: html_document
---

## Housekeeping
Preapre environment for analysis
```{r, message=F, warning=F}
library(randomForest)
library(caret)
library(rpart)
library(ElemStatLearn)
set.seed(1972)
getwd()
```

## Data Processing and Cleaning
Read data files and remove IDs and classification column for testing data
```{r, message=F, warning=F}
m.training <- read.csv("pml-training.csv", header=TRUE, sep=",", na.strings=c("NA",""))
m.training<-m.training[,-1]
m.testing <- read.csv("pml-testing.csv", header=TRUE, sep=",", na.strings=c("NA",""))
m.testing<-m.testing[,-1]
m.testing <- m.testing[,-ncol(m.testing)]
```

Check number of columns with NAS
```{r, message=F, warning=F}
tr.NAs <- numeric()
for (i in 1:ncol(m.training)) {
  tr.NAs<-c(tr.NAs,sum(is.na(m.training[,i])))
  }
tr.NAs
te.NAs <- numeric()
for (i in 1:ncol(m.testing)) {
  te.NAs<-c(te.NAs,sum(is.na(m.testing[,i])))
  }
te.NAs
```

Remove columns with NAs
```{r, message=F, warning=F}
woNAs.tr<-c(tr.NAs==0)
woNAs.te<-c(te.NAs==0)
m.training<-m.training[,woNAs.tr]
m.testing<-m.testing[,woNAs.te]
```

Create training and validating sets
```{r, message=F, warning=F}
inTrain = createDataPartition(m.training$classe, p=0.60, list=FALSE)
m.training<-m.training[inTrain,]
m.validating<-m.training[-inTrain,]
```

Run cross-validation. Calculate The Best Tune of number of variable randomly sampled.
```{r, message=F, warning=F}
fitControl = trainControl( method = "cv", number = 2)
cv <- train(classe ~ ., data = m.training, method = "rf", trControl = fitControl)
cv$bestTune$mtry
```

## Build random forest model
```{r, message=F, warning=F}
mRF.model <- randomForest(classe~.,data=m.training,  mtry = cv$bestTune$mtry)
print(mRF.model)
```

The OOB estimate of error rate is 0.12%. Check it on validating data
```{r, message=F, warning=F}
confusionMatrix(predict(mRF.model,newdata=m.validating[,-ncol(m.validating)]), m.validating$classe)
acc.val<-c(as.numeric(predict(mRF.model,newdata=m.validating[,-ncol(m.validating)])==m.validating$classe))
acc.val<-sum(acc.val)*100/nrow(m.validating)
acc.val
```

## Predict testing data for submission (+ fix difference in data type)
```{r, message=F, warning=F}
m.testing <- rbind(m.training[100, -59] , m.testing)
ForSubmission = predict(mRF.model, newdata = m.testing[-1,])
ForSubmission
```

## Prepare Files for Submission
```{r, message=F, warning=F}
pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_", i ,".txt")
    write.table(x[i], file = filename, quote = FALSE,
                row.names = FALSE, col.names = FALSE)
  }
}
pml_write_files(ForSubmission)
```

DONE

